---
title: 'RFI time course: perception task'
author: "Jeff Peterson"
date: "March 14, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r Global Options, message=FALSE, warning=FALSE}
rm(list=ls())

library(tidyverse); library(ggthemes)

source("psy_functions_source.R")

```


```{r data agg loop}

# data_directory <- "./Data/Perception_task/Good" # specify location of data
# 
# files <- list.files(data_directory) 
# 
# data_raw <- NULL # create place for data to go
# 
# for(i in 1:length(files)){
#   
#   # list files w/i subject folder
#   subfiles <- list.files(paste0(data_directory, "/", files[i])) 
#   
#   # which file in list contains "RESULTS" in its name?
#   resultsFile <- subfiles[which(str_detect(subfiles, "RESULTS") == TRUE)] 
#   
#   # load results file
#   data_raw <- import(paste0(data_directory, "/", files[i], "/", resultsFile),
#                      na.strings = c("NULL")) %>% 
#     
#     mutate(sid = files[i]) %>%  # add participant ID column
#     
#     rbind(data_raw)             # aggregate data
#   
# }

```

```{r import data}
data_raw <- rio::import("./RFI_timecourse_rawdata.csv")
```


```{r housekeeping}

data_clean <- data_raw %>%
  
  janitor::clean_names() %>%        # make colnames lowercase
  filter(trial_type == "real") %>%  # filter out practice trials
  select(sid, everything()) %>%     # make sid the first column
  arrange(sid) %>%                  # sort rows by sid
  
  mutate( # add useful columns
    
    # convert frames to milliseconds
    probe_del_ms = round(probe_del * (16 + 2/3), 0), 
    
    # recode frame tilt value
    frame_orient = ifelse(frame_tilt == "-15", "Left", "Right"),
    
    # recode staircase approach
    stair_approach = ifelse(substr(trial_variant,2,2) == "L", "Left", "Right")
    
  )

sids <- unique(data_clean$sid) # list of subject IDs

```    


First determine the PSE (point of subjective equality) at each SOA. This will tell us the orientation at which the participant would be equally likely to report the rod as being tilted to the left or right of vertical. Next, determine the magnitude of the rod-and-frame effect (RFE; also referred to as rod-and-frame illusion or RFI). For each participant at each SOA, the RFE is determined by subracting the PSE in the left-tilted frame condition from the PSE in the right-tilted frame condition.   

It is predicted that the RFE will become progressively larger as the SOA between the rod and frame decreases. The rationale being that the percieved orientation of the rod will be more influenced by the tilted frame (as reflected by changes in the PSE) when the rod and frame are presented in closer temporal proximity. Furthermore, it is predicted that the RFE will reach a plateau when the rod and frame are presented simultaneously (i.e., SOA = 0) and remain at this maximum for rods presented after frame onset (e.g., SOA = +200). 



```{r fit to raw data}

# Fit 2 parameter function to participant responses (0 or 1) at each
# SOA and frame tilt condition

x_plotlims2 <- -30:30

raw_fits <- data_clean %>% 
  group_by(sid, frame_orient, probe_del_ms) %>%
  nest() %>% 
  mutate(psy_fit = map(data, ~optim(par = c(1, 1), 
                                    psy_fun, 
                                    x_val = .$rod_tilt,
                                    rsp = .$response)),
         
         # Pull out relevant parameters from full solution
         pse = map_dbl(psy_fit, ~.$par[1]),
         
         slope = map_dbl(psy_fit, ~.$par[2]),
         
         residual = map_dbl(psy_fit, "value"),
         
         curve_fit = map(psy_fit, ~psy_plot(x_val = x_plotlims2,
                                            c(.$par[1], .$par[2])))
         
  )

```



```{r PSE calc & plot}

data_pse <- data_clean %>%
  
  # select only trials that are reversals and a reversal value greater than 6
  filter(reversal == 1, 
         reversal_num > 6) %>%
  
  # group data by...
  group_by(sid, frame_orient, probe_del_ms) %>% 
  
  # ... and find average rod tilt within each group
  summarise(pse = mean(rod_tilt))

```


```{r psy_fun}
# Calculate RFI for each participant at each SOA. Then fit sigmoid to RFEs

x_plotlims <- -400:400  # x-vals for plotting curve fit

SOAs <- sort(unique(data_pse$probe_del_ms))

data_rfe <- data_pse %>% 
  spread(frame_orient, pse) %>%       # frame orient to 2 columns
  mutate(rfe = Right - Left) %>%      # determine RFI magnitude  
  select(sid, probe_del_ms, rfe) %>%  # select columns
  nest() %>% # nest SOA and rfe for each participant into list column called "data"
  
  # Optimize psy_fun using nested data for each participant
  mutate(psy_fit = map(data, ~optim(par = c(-100, 50, 6),   # set initial parameters to fit
                                    psy_fun,                # function to minimize
                                    x_val = .$probe_del_ms, # SOA value
                                    rsp = .$rfe)),          # participant response
         
         # Pull out relevant parameters from psy_fit for easy access
         # par[1], [2], [3], and "value" are outputs from optim()
         pse = map_dbl(psy_fit, ~.$par[1]),
         
         slope = map_dbl(psy_fit, ~.$par[2]),
         
         height = map_dbl(psy_fit, ~.$par[3]),
         
         residual = map_dbl(psy_fit, "value"),
         
         # psy_plot returns x & y values of individual fitted functions
         curve_fit = map(psy_fit, ~psy_plot(x_val = x_plotlims,
                                            c(.$par[1], .$par[2], .$par[3]))),
         
         # predicted response at stimulus level
         predicted_y = map(psy_fit, ~psy_plot(x_val = SOAs,
                                              c(.$par[1], .$par[2], .$par[3]))),
         
         # Where function crosses 10% threshold... just 'cause
         thresh10 = find_thresh(0.1, c(pse, slope, height))
         
  )

```

```{r prediction error}
# Join observed and predicted errors into one data set
data_error <- data_rfe %>% 
  select(sid, data, predicted_y) %>% 
  unnest() %>% 
  rename(predicted = y_val,
         observed = rfe) %>% 
  mutate(pred_error = predicted - observed)



# define function to grab p value from cor.test
r_sq <- function(observed, predicted){
  cor.obj <- cor.test(x = observed, y = predicted)
  p_val <- cor.obj$p.val
  return(p_val)
}

# get r^2 and p value for each fit
data_rsq <- data_error %>% 
  group_by(sid) %>% 
  summarize(p_val = r_sq(observed, predicted),
            rsq = cor(observed, predicted)^2)

# which subjects had a fit with a p value greater than 0.1?
sid_rm <- unique(data_rsq$sid[which(data_rsq$p_val > 0.1)])

if(length(sid_rm) == 0){sid_rm <- "None"}

print(paste("Subjects Removed:", sid_rm))

```


```{r fitted group RFE data}
# Below follows same logic as above chunk, but applied to average RFE at each SOA instead of individual RFEs.

RFE_group_fit <- data_rfe %>% 
  select(sid, data) %>% 
  unnest() %>% 
  group_by(probe_del_ms) %>% 
  summarize(mean_rfe = mean(rfe)) %>%
  nest() %>% 
  
  mutate(psy_fit = map(data, ~optim(par = c(-100, 50, 6), 
                                    psy_fun, 
                                    x_val = .$probe_del_ms,
                                    rsp = .$mean_rfe)),
         
         # Pull out relevant parameters from full solution
         pse = map_dbl(psy_fit, ~.$par[1]),
         
         slope = map_dbl(psy_fit, ~.$par[2]),
         
         height = map_dbl(psy_fit, ~.$par[3]),
         
         residual = map_dbl(psy_fit, "value"),
         
         curve_fit = map(psy_fit, ~psy_plot(x_val = x_plotlims,
                                            c(.$par[1], .$par[2], .$par[3]))),
         
         # Find where function crosses 10% threshold   
         thresh10 = find_thresh(0.1, c(pse, slope, height))
         
  )


```



Plots

```{r graphics tuner}

# Easy way to figure themes 

line_size = 2.3
leftFrame_col = "royalblue"
rightFrame_col = "firebrick2"
upFrame_col = "grey24"
text_size = 22 
colour <- c("royalblue", "firebrick2")

```

```{r PSE plot}

plot_LR_frames <- data_pse %>%
  
  # group data by...
  group_by(frame_orient, probe_del_ms) %>%
  
  # ... get mean and standard error of each group
  summarize(mean_pse = mean(pse),
            se_pse = sd(pse) / sqrt(length(sids))) %>% 
  
  # pipe above data into ggplot
  ggplot(aes(x = probe_del_ms, y = mean_pse, color = frame_orient)) +
  geom_vline(xintercept = 0, col = "grey29", linetype = 3, size = line_size) +
  geom_line(size = line_size) +
  geom_errorbar(aes(ymin = mean_pse - se_pse, ymax = mean_pse + se_pse),
                width = 1.4, size = 1.4, 
                position = "dodge") + 
  
  geom_point(shape = 21, size = 6, fill = "white") +
  
  #scale_y_continuous(breaks = seq(-4, 3, by = 1)) +
  ylab(expression(paste("PSE (",degree,")"))) +
  #xlab("\n Probe Lead                                      Probe Lag \n SOA (msec)") +
  
  theme_bw() +
  
  theme(axis.title = element_text(size = text_size))  +
  theme(axis.title.x = element_blank()) +
  theme(axis.text = element_text(size = text_size)) +
  theme(legend.position = "none") +
  scale_colour_manual(values=colour) 

plot_LR_frames


# Copy below into console for saving image

# jpeg("lr_frames.jpg", quality = 100,
#      width = 9.2, height = 6.4, units = 'in', res = 300)
# plot_LR_frames # Make plot
# dev.off()

```



```{r RFE plot}

# Get data points for mean RFE at each SOA and the SE
grp_pts <- data_rfe %>% 
  select(sid, data) %>% 
  unnest() %>% 
  group_by(probe_del_ms) %>% 
  summarize(mean_rfe = mean(rfe),
            se_rfe = sd(rfe) / sqrt(length(sids))
  )

# Get data for smooth line
grp_fit <- RFE_group_fit %>% 
  select(curve_fit) %>% 
  unnest()


plot_TC_fitted <- ggplot(data = grp_pts, aes(x = probe_del_ms, y = mean_rfe)) +
  geom_vline(xintercept = 0, col = "grey29", linetype = 3, size = line_size) +
  geom_errorbar(aes(ymin = mean_rfe - se_rfe, ymax = mean_rfe + se_rfe),
                width = 1.4, size = 1.2, 
                position = "dodge") +
  geom_line(data = grp_fit, aes(x = x_val, y = y_val), 
            col = rightFrame_col,
            size = line_size) +
  
  geom_point(shape = 21, size = 6, fill = "white") +
  
  # scale_y_continuous(breaks = seq(-4, 3, by = 1)) +
  ylab(expression(paste("RFI Magnitude (",degree,")"))) +
  # xlab("\n Probe Lead                                      Probe Lag \n SOA (msec)") +
  
  theme_bw() +
  
  theme(axis.title = element_text(size = text_size))  +
  theme(axis.title.x = element_blank()) +
  theme(axis.text = element_text(size = text_size)) +
  #theme(legend.position = "none") +
  scale_colour_manual(values=colour) +
  xlim(-300, 300)

plot_TC_fitted


# Copy into console for saving image
# jpeg("rfi_tc.jpg", quality = 100,
#      width = 9.2, height = 6.4, units = 'in', res = 300)
# plot_TC_fitted # Make plot
# dev.off()



```



```{r individual TC overlay}

data_rfe %>% 
  select(sid, data) %>% 
  unnest() %>% 
  
  ggplot(aes(x = probe_del_ms, y = rfe, color = sid)) +          
  geom_vline(xintercept = 0,         
             col = "grey29", linetype = 3, size = line_size) +  
  
  geom_line(size = 1) +                     
  geom_point(shape = 21, size = 4, fill = "white") + 
  theme_bw() +                                       
  
  
  ylab(expression(paste("RFE (",degree,")"))) +                     
  xlab("Probe - Frame SOA")



```


```{r individual plot loop, echo=FALSE, fig.height=5, fig.width = 5, fig.show='hold', fig.align='center'}

# Create plots for each participant
# 1) staircases, 2) fits to raw data, 3) RFE time course plot



for(i in 1:length(sids)){

  
  sbj_plt_name <- sids[i]
  if(sids[i] %in% sid_rm == TRUE){
    sbj_plt_name <- paste("REMOVED", sids[i]) # note whether subject was bad
  }
  
  # Staircase plot
  plt1 <- data_clean %>%
    filter(sid == sids[i]) %>%
    
    ggplot(aes(x = trial_iteration_val, y = rod_tilt)) +
    geom_line(aes(color = stair_approach), size = 1.3) +
    facet_grid(probe_del_ms ~ frame_orient) +
    ggtitle(sids[i])
  
  # data points for raw data fit
  indv_pts <- raw_fits %>%
    filter(sid == sids[i]) %>% 
    select(frame_orient, probe_del_ms, data) %>% 
    unnest() %>%
    select(frame_orient, probe_del_ms, rod_tilt, response) %>% 
    group_by(frame_orient, probe_del_ms, rod_tilt) %>% 
    summarize(m_resp = mean(response),
              n = n())
  
  # plot fit to raw data
  plt2 <- raw_fits %>% 
    filter(sid == sids[i]) %>% 
    select(frame_orient, probe_del_ms, curve_fit) %>% 
    unnest() %>% 
    
    ggplot(aes(x = x_val, y = y_val)) + 
    geom_line() +
    geom_point(data = indv_pts, aes(x = rod_tilt, y = m_resp, size = n),
               color = rightFrame_col, alpha = .6) +
    facet_grid(probe_del_ms ~ frame_orient) +
    
    xlab("Rod Tilt") +
    ylab("Proportion Respond Rightward") +
    
    ggtitle(sbj_plt_name)
  
  # data points for time course plot
  indv_pts2 <- data_rfe %>%
    filter(sid == sids[i]) %>% 
    select(data) %>% 
    unnest() 
  
  # plot RFI time course
  plt3 <- data_rfe %>% 
    filter(sid == sids[i]) %>% 
    select(curve_fit) %>% 
    unnest() %>% 
    
    ggplot(aes(x = x_val, y = y_val)) +
    geom_vline(xintercept = 0, linetype = "dashed", size = 1) +
    geom_line(color = "firebrick2", size = line_size) +
    geom_point(data = indv_pts2, 
               aes(x = probe_del_ms, y = rfe),
               size = 3, fill = "white") +
    xlab("Probe - Frame SOA") +
    ylab(expression(paste("RFI Magnitude (",degree,")"))) +
    
    theme_bw() +
    
    ggtitle(paste0(sbj_plt_name, ": RFI timecourse"))
  
  
  # Error plot
  indv_rsq <- data_rsq %>% 
    filter(sid == sids[i]) 
  
  indv_error <- data_error %>% 
    filter(sid == sids[i]) %>% 
    
    ggplot(aes(x = observed, y = predicted)) +
    geom_abline(slope = 1, intercept = 0, size = 1, color = "firebrick2") +
    geom_point(size = 2) +
    geom_smooth(method='lm',formula=y~x) +
    # theme(axis.text = element_blank()) +
    ggtitle(paste0(sbj_plt_name, ": Prediction Error")) +
    # labs(subtitle = "Red line is unity") +
    
    geom_text(data = indv_rsq,
              mapping = aes(x = -Inf, y = Inf, label = paste("p =", round(p_val, 3))),
              hjust   = -.5,
              vjust   = 1.3) +
    geom_text(data = indv_rsq,
              mapping = aes(x = -Inf, y = Inf, label = paste("r^2 =", round(rsq, 3))),
              hjust   = -.4,
              vjust   = 2.6)
  
  
  print(plt1)
  print(plt2)
  print(plt3)
  print(indv_error)  
  
}

```


With the given data, the PSE can be estimated in two ways:
1) By determining the average rod tilt of the last six reversals
2) By fitting psychometric functions to the raw responses

Let's compare

```{r pse compare}

pse_compare <- raw_fits %>% 
  select(sid, frame_orient, probe_del_ms, pse) %>% 
  rename(pse_raw = pse) %>% 
  
  right_join(data_pse, by = c("sid", "frame_orient", "probe_del_ms")) %>% 
  mutate(pse_diff = pse - pse_raw)



pse_compare2 <- pse_compare %>% 
  group_by(frame_orient, probe_del_ms) %>% 
  summarize(mean_pse = mean(pse),
            mean_pse_raw = mean(pse_raw),
            mean_pse_diff = mean(pse_diff))

mean(pse_compare$pse_diff)
sd(pse_compare$pse_diff)

plot(pse_compare$pse_diff)

```

